---
title: "Aides Drug"
author: "Bella Verdin"
date: "Spring 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods

Variables we use are:

Treatment - categorical.   
DrugUser - categorical.    
Infected - categorical.   

Infected depends on treatment among drug users.

We get just the drug users from the population.

```{r}
drugUsersData = select(triUmp, treatment, DrugUser, infected)
drugUsersData = filter(drugUsersData, DrugUser== "yes")
```


# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->


```{r}
barchartGC(~infected + treatment,data=drugUsersData)
barchartGC(~infected + treatment,data=drugUsersData, type="percent")
```

The percent bar chart shows that those that have been given the placebo have a higher percentage, vs those who got the real drug have a lower percentage, meaning the drug is working!

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=drugUsersData)
rowPerc(table1)
colPerc(table1)
```

In the top graph those with who got aids, 24.29% of them are getting the drug and 75.71% were getting the fake pill. if the drug showed ot effective then the numbers would be around 50% & 50%, seeing as they are not then the drug might not be working .

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


In these tests I can see that the p-value is 0.0000000423. Which means that the preventative aids drug seems effective . 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

I am trying to see if this new preventative drug is effective. the p-value bing 0.0000000423 which is less than 0.05 so I will reject the null hypothesis. the drug is effective, and when looking at the bar chart it shows the preventative aid drug may be effective. In odds ratio and by taking the drug the odds are .35 to 1 that it would cure you from aids when using the drug. 


